Manuscript

### 2015-01-22 inst/simulation/manuscript.Rmd
...

context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
scenarios <- c("A","B","C","D","E","F")
modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"), paste("SmoothModel", "nbinomial", "ar1", sep = "-"), "FMPModel")
iterations <- as.integer(c(1,1,1,1,3,1))
populationSizeOverEstimate <- 1e999

@@

@@

```{r figure-1}
#window <- c(3.375, 3.4, 6.9, 6.925)*1e6
window <- (c(3.375, 3.4, 6.9, 6.925)+c(-.01,.01,-.01,.01)*3)*1e6
habitatRaster <- crop(study$studyArea$habitat, extent(window))
z <- sampleRegular(habitatRaster, 10000, asRaster=TRUE)
habitatData <- as.data.frame(rasterToPoints(z))
#habitatData <- as.data.frame(rasterToPoints(habitatRaster))
colnames(habitatData) <- c("x","y","z")
habitatData$scenario <- factor("E", levels=scenarios)
habitatData2 <- habitatData
habitatData2$scenario <- factor("F", levels=scenarios)
habitatData <- rbind(habitatData, habitatData2)

@@

p <- ggplot(habitatData) + geom_raster(aes(x, y, fill=z), alpha=.3) + scale_fill_manual(values=col) +
  geom_path(data=trackData, aes(long, lat, group=group, colour=observation), alpha=1, size=.3) +#, alpha=observation)) +
  scale_colour_manual(values=c("darkgray","black")) + # scale_alpha_manual(values=c(.5,1)) + # gray40
  geom_path(data=surveyRoutesData, aes(long, lat, group=group), size=.3, colour="blue", alpha=.5) +
  coord_fixed(ratio=1, xlim=window[1:2], ylim=window[3:4]) +
  facet_wrap(~scenario) + #coord_equal()
  theme_raster(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
#print(p)
saveFigure(p, "manuscript-figure-1.png", width=8, height=4)

@@

#p <- ggplot(boundaryDF) + geom_polygon(aes(long, lat, group=group), fill=terrain.colors(100)[1]) +
p <- ggplot(densitiesScaled) + geom_raster(aes_string(x="x", y="y", fill=yearLayer)) +
  scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent", guide=guide_legend(title=expression(paste("Density ", individuals/km^2)))) +
  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black"), guide=FALSE) + 
  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
  facet_grid(~scenario) + coord_equal() + theme_raster(base_size=10) + theme(legend.position="bottom")
print(p)
saveFigure(p, "manuscript-figure-2a.png", width=8)

@@

```{r table-1abc}
getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, full=F, file) {
  library(plyr)
  x <- ddply(populationSizes, .(Scenario, modelName), function(x) {
    if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, R2=NA))
    y <- lm(Estimated~Observed, x)
    return(data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2)))
  })
  y <- unique(meanPopulationSizes)
  y$Scenario <- y$scenario
  y$EstDev <- y$Estimated / y$True
  z <- plyr::join(x, y)
  z$Model <- prettyModelName(z$modelName)
  z$Scenario <- substr(z$Scenario, 1, 1)
  if (full) return(z)

  z <- z[,c("Model","Scenario","Slope","EstDev","R2")]
  z$EstDev <- round(z$EstDev, 2)
  if (!missing(file)) write.csv(z, file=file)
  return(z)
}

dir <- normalizePath(context$figuresDirectory)
getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, file=file.path(dir, "table-1a.csv"))
getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, file=file.path(dir, "table-1b.csv"))
getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, file=file.path(dir, "table-1c.csv"))
```

@@

  p <- ggplot(cip2, aes(scenario, Proportion, shape=Interval, colour=modelName)) +
    geom_hline(aes(yintercept=.5), color="grey90") +
    geom_hline(aes(yintercept=.95), color="grey90") +
    geom_point() +
    scale_shape_manual(values=c(16, 1)) + scale_colour_manual("Model", values=c("darkgray", "black"), labels=c("T", "ST")) +
    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
    xlab("Scenario") + ylab("Proportion within CI") +
    facet_wrap(~panel) + ylim(0, 1) +
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank())
saveFigure(p, "manuscript-figure-5.png", width=6, height=2)

@@

```{r figure-3c123, eval=FALSE}
```

@@

```{r figure-3d123, eval=FALSE}
```

@@

```{r table-2, eval=FALSE}
# TODO: Combine with table 1
study <- getStudy(scenario="A")
validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
failedEstimations <- validation$countFailedEstimations(modelNames=modelNames)
failedEstimations$pFailed <- (1 - failedEstimations$nEstimated / 50) * 100
failedEstimations
```

@@

```{r figure-4a}
study <- getStudy(scenario="A")
validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
populationCorrelations <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames[1])

@@

```{r figure-4b}
study <- getStudy(scenario="A")
validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1])

@@

```{r figure-4c}
study <- getStudy(scenario="A")
validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1])

>>>>

context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
scenarios <- c("A","B","C","D","E","F")
allScenarios <- c(scenarios, paste0(scenarios, "combined"), paste0(scenarios, "10days"))
modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"), paste("SmoothModel", "nbinomial", "ar1", sep = "-"), "FMPModel")
iterations <- as.integer(c(1,1,1,1,3,1))
populationSizeOverEstimate <- 1e999

@@

```{r figures}
study <- getMSS(scenario=scenarios[1])$study
study$studyArea$boundary@polygons[[1]]@area / 1000^2 # Study area area
200 / (study$studyArea$boundary@polygons[[1]]@area / 1000^2) # Initial population density in Finland
intersections <- study$loadIntersections(iteration=as.integer(4))
ddply(intersections$intersections@data, .(response, year), function(x) sum(x$length) / 1000) # Total sampling effort per year (km)
#ddply(intersections$intersections@data, .(response, year), function(x) sum(x$intersections))

# Mean encounter rates
aggrates <- data.frame()
for (scenario in allScenarios) {
  study <- getMSS(scenario=scenario)$study

  rates <- data.frame()
  for (iteration in 1:50) {
    intersections <- try(study$loadIntersections(iteration=as.integer(iteration)))
    if (inherits(intersections, "try-error")) next
    x <- ddply(intersections$intersections@data, .(response, year), function(x) data.frame(rate=sum(x$intersections) / (sum(x$length)/1000) / sum(x$duration)))
    rates <- rbind(rates, x)
  }

  y <- ddply(rates, .(year), function(x, scenario) data.frame(scenario=scenario, rate.mean=mean(x$rate), rate.sd=sd(x$rate)), scenario=scenario)
  aggrates <- rbind(aggrates, y)
}

subset(aggrates, year==2001)
```

@@

```{r figure-1}
#window <- c(3.375, 3.4, 6.9, 6.925)*1e6
window <- (c(3.425, 3.45, 6.95, 6.975)+c(-.01,.01,-.01,.01)*9)*1e6
habitatRaster <- crop(study$studyArea$habitat, extent(window))
z <- sampleRegular(habitatRaster, 10000, asRaster=TRUE)
habitatData <- as.data.frame(rasterToPoints(z))
#habitatData <- as.data.frame(rasterToPoints(habitatRaster))
colnames(habitatData) <- c("x","y","z")
habitatData$scenario <- factor("E", levels=scenarios)
habitatData2 <- habitatData
habitatData2$scenario <- factor("F", levels=scenarios)
habitatData <- rbind(habitatData, habitatData2)

@@

p <- ggplot(habitatData) + geom_raster(aes(x, y, fill=z), alpha=.4) + scale_fill_manual(values=col) +
  geom_path(data=trackData, aes(long, lat, group=group, colour=observation), alpha=1, size=.3) +#, alpha=observation)) +
  scale_colour_manual(values=c("gray55","black")) + # scale_alpha_manual(values=c(.5,1)) + # gray40
  geom_path(data=surveyRoutesData, aes(long, lat, group=group), size=.3, colour="blue", alpha=.5) +
  coord_fixed(ratio=1, xlim=window[1:2], ylim=window[3:4]) +
  facet_wrap(~scenario) + #coord_equal()
  theme_raster(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
#print(p)
saveFigure(p, "manuscript-figure-1.png", width=8, height=5)

@@

#p <- ggplot(boundaryDF) + geom_polygon(aes(long, lat, group=group), fill=terrain.colors(100)[1]) +
p <- ggplot(densitiesScaled) + geom_raster(aes_string(x="x", y="y", fill=yearLayer)) +
  scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent", guide=guide_legend(title=expression(paste("Population density ", (individuals/km^2))))) +
  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
  geom_point(data=agents, aes(x, y, colour=Observed), size=1.3, colour="white") +
  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black"), guide=FALSE) +
  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
  facet_grid(~scenario) + coord_equal() + theme_raster(base_size=10) + theme(legend.position="bottom")
print(p)
saveFigure(p, "manuscript-figure-2a.png", width=8)

@@

```{r table-1abc}
getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, overEstimate, full=F, file) {
  library(plyr)
  x <- ddply(populationSizes, .(Scenario, modelName), function(x, overEstimate) {
    failed <- round((sum(x$Estimated >= overEstimate) / nrow(x)) * 100)
    if (any(abs(x$Estimated) == Inf)) return(data.frame(Intercept=NA, Slope=NA, R2=NA, Failed=failed))
    y <- lm(Estimated~Observed, x)
    return(data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2), Failed=failed))
  }, overEstimate=overEstimate)

  y <- unique(meanPopulationSizes)
  y$Scenario <- y$scenario
  y$EstDev <- y$Estimated / y$True
  z <- plyr::join(x, y)
  z$Model <- prettyModelName(z$modelName)
  z$Scenario <- substr(z$Scenario, 1, 1)
  if (full) return(z)

  z <- z[,c("Scenario","Model","Slope","R2","EstDev","Failed")]
  z$EstDev <- round(z$EstDev, 2)
  if (!missing(file)) write.csv(z, file=file)
  return(z)
}

dir <- normalizePath(context$figuresDirectory)
getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(2000+1), file=file.path(dir, "table-1a.csv"))
getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(20000+1), file=file.path(dir, "table-1b.csv"))
getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(2000+1), file=file.path(dir, "table-1c.csv"))
```

@@

@@

@@

@@

@@

```{r figure-4a}
study <- getStudy(scenario="A")
validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
populationCorrelations <- validation$getValidationSpatialPopulationSizes(modelNames=modelNames[1])
populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)

@@

```{r figure-4b}
study <- getStudy(scenario="A")
validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[1])
populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)

@@

```{r figure-4c}
study <- getStudy(scenario="A")
validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
populationCorrelations <- validation$getValidationSpatialPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[1])
populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)
